---
title: 'MVC Data Analysis '
author: "Artem Arutyunov"
output: pdf_document
urlcolor: blue
---

### Loading Libraries and Data Cleaning:
```{r message=FALSE, warning=FALSE}
#Loading the libraries 
library(tidyverse)
library(latex2exp)
library(lubridate)
library(gridExtra)

#Loading the data 
data <- read_csv('MVC.csv')

#N/A values removed for the needed variables
data <- data[complete.cases(data$ACCLASS),]
data <- data[complete.cases(data$VEHTYPE),]

#Converting from Qualitative to Quantitative 
data$ACCLASS <- ifelse(data$ACCLASS == "Fatal", 1, 0)

```

### Exploratory Data Analysis:
```{r message=FALSE}

data1<-data[-which(data$VEHTYPE == "Bicycle"), ]

#Computing the summarized statistic per 25 collision each of all-vehicles collisions.
data1His<- data1 %>%
group_by(x = ceiling(row_number()/25)) %>%
 summarize(mean = mean(ACCLASS))


#histogram created for all-vehicles collisions group.
a <- ggplot(data1His)+
  geom_histogram(aes(x= mean, y=..density..),
                 binwidth = 0.04)+
  labs(x = "Fatality Rate ",
       y = "density", title = "Fatality Rate - Non-Bicycle Vechicles Involved")+
  theme_bw()

#Computing the summarized statistic per 25 collision each 
data2 <- data[complete.cases(data$VEHTYPE),]

data2<-data2[-which(data2$VEHTYPE !="Bicycle"), ]

#Computing the summarized statistic per 25 collision each 
data2His<- data2 %>%
  group_by(x = ceiling(row_number()/25)) %>%
  summarize(mean = mean(ACCLASS))

#histogram created for bicycle involved collisions group.
b<-ggplot(data2His)+
  geom_histogram(aes(x= mean, y=..density..),
                 binwidth = 0.04)+
  labs(x = "Fatality Rate",
       y = "density", title = "Fatality Rate - Bicycle Involved")+
  theme_bw()

#Producing the histograms
grid.arrange(a,b)

#Computing the sample fatality proportion of non-bicycle collisions 
cat(mean(data1$ACCLASS))

#Computing the sample fatality proportion of bicycle involved collisions
cat(mean(data2$ACCLASS))
```
\newpage
\thispagestyle{empty}
## Appendix B - Calculations for Methoogy #1:
The sample proportion of fatal injuries of non-bicycle collisions is $\hat{p}_A = 0.129$ and sample size $n_A=13325$ with $1772$ fatal collisions and the sample proportion of fatal injuries of bicycle collisions is $\hat{p}_B = 0.055$ and sample size $n_A=722$ with $40$ fatal collisions. Since we are compairing two proportions we will state the following hypothesis: 
$$H_0: p_A - p_B = 0 \textit{   and   } H_A: p_A - p_B \neq 0 $$
Note that our sample size are large enough since $n_A\hat{p}_A = 1772 \geq 10$, $n_A(1-\hat{p}_A) = 11553\ge1 10$ and $n_B(\hat{p}_B) = 40 \geq 10$, $n_B(1-\hat{p}_B) =682 \geq 10$. 

We will have to compute the pooled proportion:
$$\hat{p} = \frac{40 + 1772}{13325+722} = 0.129$$
The observation difference between sample proportions is $$0.129-0.055 = 0.074$$
By null hypothesis, we will assume that $p_A = p_B = p = 0.129$, so under null hypothesis we can conduce that: 
$$ \hat{p}_A - \hat{p}_B \dot \sim N(0,0.129(0.871)(\frac{1}{722}+\frac{1}{13325})) =N(0,0.00016)  $$
Computing our test statistic we get that:
$$z = \frac{0.074-0}{\sqrt{0.00016}} =  5.850$$
Then computing the p-vlues we get that the the probability of $P(|z|\geq 5.850) \dot = 0$.


## Appendix C - Bootsrapping code for Methoogy #2:

```{r message=FALSE}

#Setting the indexes
index.bicycle <- data$VEHTYPE == "Bicycle"
obs.bicycle <- data$ACCLASS[index.bicycle]
obs.other  <- data$ACCLASS[!index.bicycle]


#Defining the bootstrap experiment
B <- 2000
boot.mean.diff <- c()

#Performing the sampling simulation and computing the proportion difference
for(b in 1:B){
  boot.bicycle <- sample(obs.bicycle, replace = TRUE)
  boot.other <- sample(obs.other, replace = TRUE)
  boot.mean.diff[b] <-  mean(boot.other) - mean(boot.bicycle) 
}

#Computing the confidence interval 
ci.mean.diff <- quantile(boot.mean.diff, probs = c(0.025, 0.975))
ci.mean.diff
```








